\textbf{SPECFEM3D Cartesian}\\ \textbf{User Manual}

### 2012-10-22 doc/USER_MANUAL/manual_SPECFEM3D_Cartesian.tex
added IBM BlueGene compilation instructions

Since then it has been developed and maintained by a development team: in alphabetical order,
Piero Basini,
C\'eline Blitz,
Ebru Bozda\u{g},
Emanuele Casarotti,
Joseph Charles,
Min Chen,
Vala Hj\"orleifsd\'ottir,
Sue Kientz,
Dimitri Komatitsch,
Jes\'us Labarta,
Nicolas Le Goff,
Pieyre Le Loher,
Qinya Liu,
Yang Luo,
Alessia Maggi,
Federica Magnoni,
Roland Martin,
Dennis McRitchie,
Matthias Meschede,
Peter Messmer,
David Mich\'ea,
Tarje Nissen-Meyer,
Daniel Peter,
Max Rietmann,
Brian Savage,
Bernhard Schuberth,
Anne Sieminski,
Leif Strand,
Carl Tape,
Jeroen Tromp,
Zhinan Xie,
Hejun Zhu.\\
%\toall{add full list of developers here and update the AUTHORS file accordingly}

@@

>>>>

Since then it has been developed and maintained by a development team: in alphabetical order,
Piero Basini,
C\'eline Blitz,
Ebru Bozda\u{g},
Emanuele Casarotti,
Joseph Charles,
Min Chen,
Vala Hj\"orleifsd\'ottir,
Sue Kientz,
Dimitri Komatitsch,
Jes\'us Labarta,
Nicolas Le Goff,
Pieyre Le Loher,
Qinya Liu,
Yang Luo,
Alessia Maggi,
Federica Magnoni,
Roland Martin,
Ren\'e Matzen,
Dennis McRitchie,
Matthias Meschede,
Peter Messmer,
David Mich\'ea,
Tarje Nissen-Meyer,
Daniel Peter,
Max Rietmann,
Brian Savage,
Bernhard Schuberth,
Anne Sieminski,
Leif Strand,
Carl Tape,
Jeroen Tromp,
Zhinan Xie,
Hejun Zhu.\\
%\toall{add full list of developers here and update the AUTHORS file accordingly}

@@

\section{Compiling on an IBM BlueGene}

\underline{More recent installation instruction for IBM BlueGene, from October 2012:}\\

\noindent
Edit file \texttt{flags.guess} and put this for \texttt{FLAGS\_CHECK} and \texttt{FLAGS\_NO\_CHECK}:\\

\noindent
\texttt{-g -qfullpath -O2 -qsave -qstrict -qtune=qp -qarch=qp -qcache=auto -qhalt=w}\\
\noindent
\texttt{-qfree=f90 -qsuffix=f=f90 -qlanglvl=95pure -Wl,-relax -Q -Q+rank,swap\_all}.\\

\noindent
The most relevant are the -qarch and -qtune flags, otherwise if these flags are set to ``auto'' then they are wrongly assigned to
the architecture of the frond-end node, which is different from that on the compute nodes.
You will need to set these flags to the right architecture for your BlueGene compute nodes, which is not necessarily ``qp'';
ask your system administrator.
On some machines if is necessary to use -O2 in these flags instead of -O3 due to a compiler bug of the XLF version installed.
We thus suggest to first try -O3, and then if the code does not compile or does not run fine then switch back to -O2.
The debug flags (-g, -qfullpath) do not influence performance but are useful to get at least some insights in case of problems.\\

\noindent
Before running \texttt{configure}, select the XL Fortran compiler by typing \texttt{module load bgq-xl/1.0}
or \texttt{module load bgq-xl} (another, less efficient option is to load the GNU compilers using \texttt{module load bgq-gnu/4.4.6} or similar).\\

\noindent
Then, to configure the code, for the SPECFEM3D\_GLOBE version type this:\\

\noindent
\texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}\\

\noindent
and for SPECFEM3D\_Cartesian type this:\\

\noindent
\texttt{./configure FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r LOCAL\_PATH\_IS\_ALSO\_GLOBAL=true}.\\

\noindent
\underline{Older installation instruction for IBM BlueGene, from 2011:}\\

\noindent
To compile the code on an IBM BlueGene, Laurent L\'eger from IDRIS, France, suggests the following: compile the code with

\noindent
\texttt{   FLAGS\_NO\_CHECK="-O3 -qsave -qstrict -qtune=auto -qarch=450d -qcache=auto \\}
\texttt{   -qfree=f90 -qsuffix=f=f90 -g -qlanglvl=95pure -qhalt=w -Q -Q+rank,swap\_all -Wl,-relax"}

\noindent
Option "-Wl,-relax" must be added on many (but not all) BlueGene systems to be able to link the binaries \texttt{xmeshfem3D}
and \texttt{xspecfem3D} because the final link step is done by the GNU \texttt{ld} linker even if
one uses \texttt{FC=bgxlf90\_r, MPIFC=mpixlf90\_r} and \texttt{CC=bgxlc\_r} to create all the object files.
On the contrary, on some BlueGene systems that use the native AIX linker option "-Wl,-relax" can lead to problems and must be suppressed from \texttt{flags.guess}.

\noindent
Also, \texttt{AR=ar, ARFLAGS=cru} and \texttt{RANLIB=ranlib} are hardwired in all \texttt{Makefile.in} files by default, but to cross-compile on BlueGene/P one needs to change
these values to \texttt{AR=bgar, ARFLAGS=cru} and \texttt{RANLIB=bgranlib}. Thus the easiest thing to do is to modify
all \texttt{Makefile.in} files and the \texttt{configure} script to set them automatically by \texttt{configure}.
One then just needs to pass the right commands to the \texttt{configure} script:

\noindent
\texttt{   ./configure --prefix=/path/to/SPECFEM3DG\_SP --host=Babel --build=BGP \\}
\texttt{      FC=bgxlf90\_r MPIFC=mpixlf90\_r CC=bgxlc\_r AR=bgar ARFLAGS=cru \\}
\texttt{      RANLIB=bgranlib LOCAL\_PATH\_IS\_ALSO\_GLOBAL=false}

\noindent
This trick can be useful for all hosts on which one needs to cross-compile.

\noindent
On BlueGene, one also needs to run the \texttt{xcreate\_header\_file} binary file manually rather than in the Makefile:

\noindent
   \texttt{bgrun -np 1 -mode VN -exe ./bin/xcreate\_header\_file}
